
This chapter introduces the purpose and overall direction 
of the thesis. It presents the motivation for the study 
and describes the main problem that is addressed. 
The research questions are then defined to guide the work. 
Finally, the chapter provides an overview 
of the structure of the thesis.

\section{Motivation}

Machine learning has seen strong growth in recent years 
and is now widely applied in product development 
and production. It is used for tasks such as optimization, 
control, prediction, and decision support 
\parencite{wuest2016ml_manufacturing}. 

Advances in data availability and computing power 
have further increased the use of machine learning 
in industrial settings \parencite{rai2021ml_manufacturing_industry4}. 
As a result, machine learning plays an important role 
in improving efficiency, supporting decision-making, 
and enabling new capabilities in engineering and manufacturing.

At the same time, the field of machine learning 
is large and diverse. Many different algorithms, 
methods, and approaches are available, 
each with its own strengths and weaknesses 
\parencite{wuest2016ml_manufacturing}. 
There is no single method that performs best 
for all problems, and performance depends on 
the specific application and data 
\parencite{adekoya2022applications_ai_em,lee2020machine_learning_enterprises}. 

This creates a challenge for engineers and practitioners, 
who must select suitable methods for their problems. 
The large number of available options can act as a barrier 
and limit the effective use of machine learning in practice 
\parencite{adekoya2022applications_ai_em}. 
Selecting an appropriate method often requires balancing 
multiple factors such as accuracy, interpretability, 
and data requirements 
\parencite{plathottam2023review_ai_manufacturing,ali2017accurate_multi_criteria_decision_making_methodology_recommending_machine_learning_algorithm}. 

If the selection is not done carefully, the resulting model 
may perform poorly or fail to provide useful insights 
\parencite{ali2017accurate_multi_criteria_decision_making_methodology_recommending_machine_learning_algorithm}. 
In practice, simpler models are often preferred 
due to their interpretability, even if more complex models 
may offer higher accuracy \parencite{paleyes2022challenges_deploying_ml_survey_case_studies}. 
This highlights the importance of considering trade-offs 
in real-world applications.

Another challenge is the gap between research 
and practical implementation. Research results are 
distributed across many articles and domains, 
which makes it difficult to gain a complete overview 
\parencite{adekoya2022applications_ai_em}. 
In addition, most studies focus on developing models, 
while less attention is given to how they should be selected 
and applied in practice.

Companies also face challenges related to lack of knowledge 
and experience when adopting machine learning 
\parencite{sonntag2023pattern_ml_product_development}. 
In many cases, the selection of methods is based on 
trial and error, which is time-consuming and inefficient 
\parencite{sonntag2023pattern_ml_product_development}. 
There is also often pressure to adopt machine learning 
without a clear plan, which increases the risk 
of unsuccessful implementations 
\parencite{plathottam2023review_ai_manufacturing}. 

These challenges highlight the need for more structured 
approaches that can support method selection 
and make machine learning more accessible.

% \begin{itemize}
%     \item Increasing use of machine learning in product development and production
%     \item Difficulty in selecting suitable ML methods for specific problems
%     \item Gap between research literature and practical implementation
%     \item Need for structured decision support
%     \item Potential value of combining ontology and literature-based retrieval
%     \item Relevance for engineers and practitioners with limited ML expertise
% \end{itemize}

\section{Problem Description}

Selecting appropriate machine learning methods 
for real-world problems is a complex task 
that depends on several interacting factors. 
These include the characteristics of the data, 
the problem context, and the requirements 
for performance and interpretability.

In practice, there is no single algorithm 
that performs well for all problems, 
and the performance of a method may vary 
across different applications and datasets 
\parencite{adekoya2022applications_ai_em,lee2020machine_learning_enterprises}. 
This means that method selection cannot be based 
on fixed rules, but must consider the specific 
conditions of each problem.

A key difficulty is that there is often no clear 
relationship between the available data, 
the problem setting, and the choice of method. 
Studies show that there is limited correlation 
between data characteristics, selected algorithms, 
and the problems they are applied to 
\parencite{weichert2019review_ml_optimization_production}. 
In addition, the success of a method depends strongly 
on the structure and quality of the data, 
which is not always known in advance 
\parencite{dogan2021machine_learning_data_mining_manufacturing}.

As a result, method selection is frequently based 
on testing multiple algorithms in practice 
\parencite{wuest2016ml_manufacturing}. 
This process is time-consuming and does not guarantee 
that the most suitable method is identified. 
It also requires experience, which many practitioners 
do not have \parencite{sonntag2023pattern_ml_product_development}.

Another challenge is that method selection often involves 
balancing several criteria. These include accuracy, 
interpretability, computational cost, and data requirements. 
Focusing on a single metric, such as accuracy, 
can lead to suboptimal choices 
\parencite{ali2017accurate_multi_criteria_decision_making_methodology_recommending_machine_learning_algorithm,sala2018selecting_ml_algorithm}. 
In real-world applications, simpler and more interpretable 
models are often preferred, even if they are less accurate 
\parencite{paleyes2022challenges_deploying_ml_survey_case_studies}. 
This reflects the importance of considering trade-offs 
in practical settings.

Interpretability itself is a critical factor. 
Many machine learning models act as black boxes, 
which makes it difficult to understand 
how results are produced 
\parencite{presciuttini2024ml_iot_manufacturing_operations}. 
This can reduce trust and limit adoption, 
especially in industrial environments 
where explanations are required.

Another limitation is the lack of structured support 
for method selection. Existing approaches 
often rely on general guidelines or simplified criteria, 
which may not capture the complexity of real problems 
\parencite{sala2018selecting_ml_algorithm}. 
In addition, most research focuses on developing models, 
while less attention is given to how they should be selected 
and applied in specific contexts 
\parencite{arena2024conceptual_framework_ml_algorithm_selection_predictive_maintenance}.

These challenges highlight the need for more systematic 
approaches to method selection. In particular, there is a need 
for solutions that can connect problem descriptions, machine 
learning methods, and supporting research. 

The large and diverse body of machine learning literature 
contains valuable knowledge, but it is difficult to use in practice. 
Combining structured knowledge with information from this literature 
can support more informed method selection and early-stage implementation.

There is therefore a need for tools that can:

\begin{itemize}
    \item support method selection based on problem descriptions
    \item provide interpretable and transparent recommendations
    \item link methods to relevant research literature
    \item support early-stage implementation
\end{itemize}
% \begin{itemize}
%     \item Users struggle to identify relevant ML methods for their problem
%     \item Existing approaches are either too generic or too technical
%     \item Lack of tools linking:
%     \begin{itemize}
%         \item problem descriptions
%         \item ML methods
%         \item supporting research articles
%     \end{itemize}
%     \item Challenge of utilizing large volumes of literature effectively
%     \item Need for:
%     \begin{itemize}
%         \item interpretable recommendations
%         \item traceability to literature
%         \item support for early implementation steps
%     \end{itemize}
%     \item Context: system combining ontology + embeddings + article retrieval
% \end{itemize}

\section{Research Questions}

Based on the challenges described above, 
this thesis addresses the following research questions:

\begin{itemize}
    \item How can ontology-based knowledge improve machine learning method selection?
    \item How can semantic similarity between problem descriptions and literature support recommendations?
    \item How can structured knowledge and unstructured text be combined in a single system?
    \item To what extent can such a system support early-stage machine learning implementation?
\end{itemize}

\section{Thesis Structure}

The remainder of this thesis is structured as follows:

\begin{itemize}
    \item Chapter 1 introduces the motivation, problem description, and research questions
    \item Chapter 2 presents background and related work
    \item Chapter 3 describes the methodology
    \item Chapter 4 presents the system implementation
    \item Chapter 5 presents the results
    \item Chapter 6 discusses the findings and limitations
    \item Chapter 7 concludes the thesis and suggests future work
\end{itemize}

% \begin{itemize}
%     \item Chapter 1: Introduction
%     \item Chapter 2: Background and state of the art 
%     \item Chapter 3: Methodology 
%     \item Chapter 4: Implementation 
%     \item Chapter 5: Results 
%     \item Chapter 6: Discussion 
%     \item Chapter 7: Conclusion and future work
% \end{itemize}




% About Pre-Master's thesis work:
% \begin{itemize}
%     \item Prior work:
%     \begin{itemize}
%         \item semi-structured analysis of a large set of articles
%         \item development of an article-based roadmap
%     \end{itemize}
%     \item Transition:
%     \begin{itemize}
%         \item need to formalize and structure knowledge
%     \end{itemize}
%     \item Current thesis:
%     \begin{itemize}
%         \item extends previous work
%         \item introduces ontology, embeddings, and system implementation
%         \item aims to automate and generalize the workflow
%     \end{itemize}
% \end{itemize}





