\section{Related Work and Existing Tools}

Few existing tools directly address the combination of 
method selection, knowledge representation, and 
implementation support that this thesis investigates. 
Most related tools either focus on automating a specific 
part of the ML workflow or provide broad access to AI 
resources without guiding users through choosing a 
suitable method for their problem. The following sections 
review the most relevant examples found.

\subsection{Ontology-Based Meta AutoML}

OMA-ML \parencite{zender2022ontology_based_meta_automl} 
is a system that combines multiple AutoML solutions 
through the ML Ontology, described in \ref{subsection:ml-ontology} 
The ML Ontology is used to filter configuration 
options and select appropriate AutoML strategies, 
making the process more transparent than standard 
AutoML systems. It also supports a wider range of 
methods by integrating solutions across multiple 
ML libraries rather than being tied to a single one.
Like other AutoML tools, OMA-ML requires the user 
to provide a labeled dataset and define the ML task 
upfront. 

\subsection{Template-Based Code Generation}

Traingenerator \parencite{rieke2021traingenerator} 
is a web-based tool that generates template Python 
code and Jupyter notebooks for ML tasks. Users select 
a task and framework from a menu and receive a working 
code template covering common steps such as data 
loading, model setup, training, and evaluation. 
The generated code can be downloaded as a Python 
script or notebook, and optionally opened directly 
in Google Colab.

The tool assumes the user has already decided on a 
method and framework before use. At the time of 
writing, the hosted version of Traingenerator 
produces a runtime error and does not render output 
correctly, likely due to an outdated Streamlit API 
call. The tool appeared to function as intended 
during the development period of this thesis.

\subsection{AI-on-Demand Platform}
The European AI-on-Demand Platform (AIoDP) and its 
DeployAI implementation offer a broad marketplace of 
pre-trained models, reusable AI modules, and 
domain-specific applications supported by cloud 
and HPC infrastructure 
\parencite{troumpoukis2025deployai_ai_on_demand_platform}. 
The platform targets SMEs, researchers, and public 
sector organizations and provides tools for building, 
training, and deploying AI solutions at scale. 
Users can browse and filter available solutions 
to narrow down options relevant to their domain 
or use case. The primary focus is on access and 
deployment of existing solutions rather than on 
guiding users through method selection for a 
new problem.